---
title: "Steps to create a new R packages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Steps to create a new R packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  dev = "png",
  dpi = 150,
  fig.asp = 1,
  fig.width = 7,
  fig.height = 5,
  out.width = "100%",
  out.height = "70%",
  fig.align = "center",
  collapse = TRUE,
  comment = "#>"
)
```

<br>

<p style="text-align:justify;"> This section shows in detail how to create a new package in the statistical software R Studio. Starting with a presentation, followed by a script in R code detailing the steps in an understandable way and finally a video tutorial on how to create a new package with simple functions and with all the suggestions given for its creation.</p>

<br>

# `r colorize("Presentation of the creation of a new R package","#1CA666")`

<object data="Tutorial/RPackage.pdf" width=775, height=550></object>

<br>

# `r colorize("Manual to create a new R package.","#1CA666")`

<object data="Tutorial/EPN_Manual_crear_paquete.pdf" width=775, height=550></object>

<br>

# `r colorize("Scrip to create a new R package.","#1CA666")`

<object data="Tutorial/ScripNuevoPaquete.pdf" width=775, height=550></object>

<br>


